---
title: "phase 1 analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readxl)
data <- read_excel("NIH P1 Emotion Understanding SDS300.xlsx")

data <- janitor::clean_names(data, 'snake')
```


Univariate Analysis:
```{r}
#males vs females in the study
ggplot(data, aes(x = gender_1)) +
  geom_histogram()

summary(data$gender_1)

#distribution of word spans according to gender
ggplot(data, aes(x = word_span)) +
  geom_histogram() +
  facet_grid(vars(gender))

summary(data$word_span)

#age distribution for males vs females
ggplot(data, aes(x = age_1)) +
  geom_histogram() +
  facet_grid(vars(gender))

summary(data$age_1)

#PLS4 score distribution male vs female
ggplot(data, aes(x = change_pls4)) +
  geom_histogram() +
  facet_grid(vars(gender)) +
  labs(x = "Change in PLS-4 Score")

summary(data$pls4comp_raw1)
summary(data$pls4_comp_raw3)
#increase from 1 to 3

#EOWPVT raw score distribution male vs female 
ggplot(data, aes(x = change_eowpvt)) +
  geom_histogram() +
  facet_grid(vars(gender)) +
  labs(x = "Change in EOWVPT Score")

ggplot(data, aes(x = change_emo_total)) +
  geom_histogram() +
  facet_grid(vars(gender)) +
  labs(x = "Change in total emotional score")

summary(data$eowpvt_raw1)
summary(data$eowpvt_raw3)
#increase from 1 to 3

#EMO sit 1 distrubution
ggplot(data, aes(x = emo_sit_1)) +
  geom_histogram() +
  facet_grid(vars(gender))

summary(data$emo_sit_1)
summary(data$emo_sit_3)
#only slight inc

#look at total pre emotional score for intervention groups vs no intervention
ggplot(data, aes(x = change_emo_total)) +
  geom_histogram() +
  facet_grid(vars(intervention))

summary((data %>% filter(intervention == 1))$change_emo_total)
summary((data %>% filter(intervention == 0))$change_emo_total)
#both groups start off pretty similar

summary((data %>% filter(intervention == 0))$change_emo_total)

```

```{r}
ggplot(data, aes(x = change_pls4)) +
  geom_histogram()
ggplot(data, aes(x = change_eowpvt)) +
  geom_histogram()
ggplot(data, aes(x = change_emo_total)) +
  geom_histogram()
```


Bivariate Analysis
```{r}
#word span vs total emotional score
ggplot(data, aes(x = word_span, y = emo_total_1)) +
  geom_point() +
  geom_jitter()

ggplot(data, aes(x = word_span, y = emo_total_3)) +
  geom_point() +
  geom_jitter()

#it looks like there is a more linear relationship for 3 than 1




#scatterplot between word span and emotion score - do kids with a higher word span have better emotional understanding? 
ggplot(data, aes(x = word_span, y = emo_total_3, color = gender)) +
  geom_point() +
  geom_jitter()
```


try different linear regression models

WORD SPAN VS EMOTION SCORE AT END OF INTERVENTION:
```{r}
#control
mod1 <- lm(emo_total_3 ~ word_span, data = data %>% filter(i_group == 1))
summary(mod1)
#very weak R and R^2, but filtering for intervention improves it

#compare to explicit:
mod1e <- lm(emo_total_3 ~ word_span, data = data %>% filter(i_group == 2))
summary(mod1e)

#compare to implicit:
mod1i <- lm(emo_total_3 ~ word_span, data = data %>% filter(i_group == 3))
summary(mod1i)
```


WORD SPAN VS PLS-4 SCORE AT END OF INTERVENTION:
```{r}
#control
mod2 <- lm(pls4_comp_raw3 ~ word_span, data = data %>% filter(i_group == 1))
summary(mod2)

#explicit emotion group
mod3 <- lm(pls4_comp_raw3 ~ word_span, data = data %>% filter(i_group == 2))
summary(mod3)

#implicit emotion group
mod4 <- lm(pls4_comp_raw3 ~ word_span, data = data%>% filter(i_group == 3))
summary(mod4)
#similar strength to control group
```

WORD SPAN + EOWPVT + PLS-4 SCORESVS EMOTIONAL SCORE AT END OF INTERVENTION:
```{r}
#control
mod5 <- lm(emo_total_3 ~ word_span + eowpvt_raw3 + pls4_comp_raw3, data = data %>% filter(i_group == 1))
summary(mod5)
#basically no linear correlation at all for control group
  
#explicit emotion group
mod6 <- lm(emo_total_3 ~ word_span + eowpvt_raw3 + pls4_comp_raw3, data = data %>% filter(i_group == 2))
summary(mod6)
#more linear correlation for intervention group!

#implicit emotion group
mod6i <- lm(emo_total_3 ~ word_span + eowpvt_raw3 + pls4_comp_raw3, data = data %>% filter(i_group == 3))
summary(mod6i)
```


NEXT: maybe try logistic regression for gender, doing models with binary variables for intervention group?



FINAL ANALYSIS

add new variables to get change in scores for each student
```{r}
data <- data %>%
  mutate(change_pls4 = (pls4_comp_raw3 - pls4comp_raw1)) %>%
  mutate(change_eowpvt = (eowpvt_raw3- eowpvt_raw1)) %>%
  mutate(change_emo_total = (emo_total_3 - emo_total_1)) %>%
  mutate(change_emo_sit = (emo_sit_3 - emo_sit_1)) %>%
  mutate(change_emo_desire = (emo_desire_3 - emo_desire_1)) 
```


multiple regression
```{r}

simple_mod <- lm(change_pls4 ~ i_group, data = data)
summary(simple_mod)

simple_mod2 <- lm(change_eowpvt ~ i_group, data = data)
summary(simple_mod2)

simple_mod3 <- lm(change_emo_total ~ i_group, data = data)
summary(simple_mod3)
```

```{r}
mult_mod_iq <- lm(change_pls4 ~ i_group + nviq_ss + (i_group*nviq_ss), data = data)
summary(mult_mod_iq)

mult_mod_iq_word <- lm(change_pls4 ~ i_group + nviq_ss + word_span + (i_group*nviq_ss) + (i_group*word_span) + (nviq_ss*word_span), data = data)
summary(mult_mod_iq_word)

mult_mod_iq_word_age <- lm(change_pls4 ~ i_group + nviq_ss + word_span + age_1 + (i_group*nviq_ss) + (i_group*word_span) + (i_group*age_1) + (nviq_ss*word_span) + (nviq_ss*age_1) + (word_span*age_1), data = data)
summary(mult_mod_iq_word_age)
```

```{r}
summary(lm(change_eowpvt ~ i_group + nviq_ss + (i_group*nviq_ss), data = data))

summary(lm(change_eowpvt ~ i_group + nviq_ss + word_span + (i_group*nviq_ss) + (i_group*word_span) + (nviq_ss*word_span), data = data))

summary(lm(change_eowpvt ~ i_group + nviq_ss + word_span + age_1 + (i_group*nviq_ss) + (i_group*word_span) + (i_group*age_1) + (nviq_ss*word_span) + (nviq_ss*age_1) + (word_span*age_1), data = data))
```

```{r}
summary(lm(change_emo_total ~ i_group + nviq_ss + (i_group*nviq_ss), data = data))

summary(lm(change_emo_total ~ i_group + nviq_ss + word_span + (i_group*nviq_ss) + (i_group*word_span) + (nviq_ss*word_span), data = data))

summary(lm(change_emo_total ~ i_group + nviq_ss + word_span + age_1 + (i_group*nviq_ss) + (i_group*word_span) + (i_group*age_1) + (nviq_ss*word_span) + (nviq_ss*age_1) + (word_span*age_1), data = data))
```

